A Short Guide to Anyons and Modular Functors

### 2016-10-11 guide.tex
more changes
some fixes from DW

\documentclass[aps, letterpaper, onecolumn, superscriptaddress, notitlepage, 11pt, groupedaddress]{revtex4-1}
\usepackage{times}

@@

The algebraic side is known to mathematicians as the study of braided
tensor categories, or more specifically, modular tensor categories.
This algebraic language appears to be more commonly used 
in the physics literature, such as the well cited Appendix E of Kitaev~\cite{Kitaev2006}.
Such algebraic calculations can be interpreted as
manipulations of string diagrams~\cite{Baez2010},
or \emph{skeins}.
These strings encode connections in the algebra, for example 
the Einstein summation convention. But they also faithfully
encode the twisting that occurs when anyons are re-ordered or
braided around each other. And these kinds of spatial relationships
are fruitfully studied using topological methods.

@@

The main goal of the present work is to sketch how a
braided tensor category arises from a modular functor.
In the mathematical literature,
this is covered in \cite{Turaev1994,Tillmann1998,Bakalov2001}
but as we just noted they use a different formulation for a
modular functor.
%The much more ambitious task of deriving a modular
%functor from a braided tensor category is

@@

In this section we begin with the familiar question 
of particle exchange in three dimensions, whose answer is bosons and fermions.
We then show how in restricting the particles to two dimensions many
more possibilities arise.
Our focus will be on the close connection between the algebraic and the topological viewpoints,
aiming to motivate the definition of a modular functor given in the next section.

@@

For our purposes a \emph{surface} will be a compact
oriented $2$-dimensional differentiable manifold with boundary. % smooth or topological ?
We will not require surfaces to be connected.
%Each boundary component is oriented (either explicitly or
%otherwise by using the orientation of the manifold)
By a \emph{hole} of $M$ we mean a component of its boundary. % $\partial M$.
Each hole will inherit the manifold orientation,
contain a distinguished \emph{base point},
and be labeled with
%an element from a finite set $\mathbb{A}=\{1, a, b, c ...\}.$
an element from a fixed finite set $\A.$
This is the set of ``anyon labels'', and comes
equipped with a vacuum element $\vac$ 
and an involution $\ \widehat{}\ $ such that $\widehat{\vac}=\vac.$
The involution maps an anyon label to the ``antiparticle'' label.

>>>>

\documentclass[aps, tightenlines, letterpaper, onecolumn, superscriptaddress, notitlepage, 11pt, groupedaddress]{revtex4-1}
\usepackage{times}

@@

The algebraic side is known to mathematicians as the study of braided
fusion tensor categories, or more specifically, modular tensor categories.
This algebraic language appears to be more commonly used 
in the physics literature, such as the well cited Appendix E of Kitaev~\cite{Kitaev2006}.
Such algebraic calculations can be interpreted as
manipulations of string diagrams~\cite{Baez2010},
or \emph{skeins}.
These strings encode connections in the algebra, for example 
the Einstein summation convention. But they also faithfully
encode the twisting that occurs when anyons are re-ordered or
braided around each other. And these kinds of spatial relationships
are fruitfully studied using topological methods.

@@

The main goal of the present work is to sketch how a
braided fusion tensor category arises from a modular functor.
In the mathematical literature,
this is covered in \cite{Turaev1994,Tillmann1998,Bakalov2001}
but as we just noted they use a different formulation for a
modular functor.
%The much more ambitious task of deriving a modular
%functor from a braided tensor category is

@@

In this section we begin with the familiar question 
of particle exchange statistics in three dimensions, whose answer is bosons and fermions.
We then show how in restricting the particles to two dimensions many
more possibilities arise.
Our focus will be on the close connection between the algebraic and the topological viewpoints,
aiming to motivate the definition of a modular functor given in the next section.

@@

For our purposes a \emph{surface} will be a compact
oriented $2$-dimensional differentiable manifold with boundary. % smooth or topological ?
We will not require surfaces to be connected.
%Each boundary component is oriented (either explicitly or
%otherwise by using the orientation of the manifold)
By a \emph{hole} of $M$ we mean a connected component of its boundary. % $\partial M$.
Each hole will inherit the manifold orientation,
contain a distinguished \emph{base point},
and be labeled with
%an element from a finite set $\mathbb{A}=\{1, a, b, c ...\}.$
an element from a fixed finite set $\A.$
This is the set of ``anyon labels'', and comes
equipped with a vacuum element $\vac$ 
and an involution $\ \widehat{}\ $ such that $\widehat{\vac}=\vac.$
The involution maps an anyon label to the ``antiparticle'' label.
